---
title: "PAW Project.Rmd"
author: "Jamin Chai"
date: "11 0ct, 2017"
output: html_document
---

## Background

Using devices such as Jawbone Up, Nike FuelBand and Fitbit, it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data Preparing

First, we load the two datasets.

```{r}
pmltraining <- read.csv('pml-training.csv')
pmltesting <- read.csv('pml-testing.csv')
```

Second, we divide training data into two sets for training and cross validation.

```{r}
library(caret)
training_parition <- createDataPartition(pmltraining$classe, list=FALSE)
training = pmltraining[training_parition,]
testing = pmltraining[-training_parition,]
```

Then, we choose columns that contains only "numeric".

```{r}
num_index = which(lapply(training, class) %in% c('numeric')  )
```

Furthermore, we prepare one pre-model and two pre-datasets for incoming modeling.

```{r}
pre_model <- preProcess(training[,num_index], method=c('knnImpute'))
pre_training <- cbind(training$classe, predict(pre_model, training[,num_index]))
pre_testing <- cbind(testing$classe, predict(pre_model, testing[,num_index]))
preTesting <- predict(pre_model, pmltesting[,num_index])
```

Meanwhile, we need to fix labels on "classe".

```{r}
names(pre_training)[1] <- 'classe'
names(pre_testing)[1] <- 'classe'
```

## Modeling

We need to use "random forest" to build the model. And we will see later that the model provides considerable accuracy to predict the 20 test cases.

```{r}
library(randomForest)
set.seed(1000)
model  <- randomForest(classe ~ ., data = pre_training, ntree=100)
```

## Cross Validation

In this stage, we need to measure the accuracy using our training set and our cross validation set.

```{r}
ex_training <- predict(model, pre_training) 
print(confusionMatrix(ex_training, pre_training$classe))
```

As is shown in the results, the model itself is not biased.

```{r}
ex_testing <- predict(model, pre_testing) 
print(confusionMatrix(ex_testing, pre_testing$classe))
```

Aa a consequencee, the model outcomes prediction accuracy greater than 99%, so that we could provide accurate prediction of the testing set.

## Results

All in all, our prediction for the testing set is displayed below.

```{r}
results <- predict(model, preTesting) 
print(results)
```
